The Essence of Summarization\\ in Pushdown Flow Analyses\\ \textbf{Supplemental Material}

### 2014-03-25 content.tex
Proof for PDCFA finished?
Fixed up (still contrived) example. Added Racket implementation. Small fix to reify and F.

%%
This paper gives a common, simple methodology to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
%
Our presentation is meant to be a tutorial to this related literature, with the additional promise of a novel analysis.
%
We don't presume a language of formal semantics such that our technique applies to all formalisms in such a language.
%
An overly formal approach is unlikely to capture all ``real language's'' formalizations.
%
The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
%
We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
%
This paper has been partially mechanized in the Coq proof assistant, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.

@@

This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
%
As it is, the CESK machine approximates the tabular machine, but not the other way around;
%
if we use a finite allocation strategy, then we don't explore some non-deterministic paths that lead to different results for a function call.
%
We thus explore all paths and share $\mktab$ and $\mmemo$ between all paths.

To turn this semantics into PDCFA's algorithm for constructing a pushdown system with only reachable states, given the appropriate $\alloc$, we apply a widening operator to make $\mktab$, and $\mmemo$ shared amongst all states, in what is then called a $\System$. We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
%

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System
\end{align*}
\begin{align*}
  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{align*}
    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge \mastate\nequiv\tpl{\mval,\mstore,\krt{\mctx}}} \\
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
  \end{align*}
}
%%

%%
A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
%
We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%
The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
%
Note the ordering on systems is pointwise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
%

\paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself, so $\menv$ is always the identity function $\iota$) to run the following example:

\begin{center}
  \todo{fix example}
  %\includegraphics[scale=0.5]{example0}
\end{center}
  % \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

\begin{figure}
      % \begin{tabular}{rcccl}
      %   $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
      %   & $\Rightarrow$ &
      %   $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
      %   & $\Rightarrow$ &
      %   $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
      % \end{tabular}
      % \begin{center}
      % \begin{tabular}{|l|c|c|c|}
      %   \hline
      %   \# & $\mktab$ & $\mmemo$ \\
      %   \hline
      %   0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
      %   1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
      %   \hline
      %   \end{tabular}
      % \begin{tabular}{|l|c|}
      %   \hline
      %   \# & $\mctx$ \\
      %   \hline
      %   0  & $(({\tt x}, \iota), \mstore_1)$ \\
      %   1  & $(({\tt x}, \iota), \mstore_3)$ \\
      %   \hline
      % \end{tabular}
      % \begin{tabular}{|l|c|c|}
      %   \hline
      %   \# & $\mstore$ & $\mkont$ \\
      %   \hline
      %   0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
      %   1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
      %   2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
      %   3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
      %   4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
      %   \hline
      % \end{tabular}
      % \end{center}
\todo{fix example}
      \caption{Example analysis evaluation}
      \label{fig:ex-eval}
\end{figure}
% %% FIXME stores in memo-table
% \begin{enumerate}
% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
% \end{enumerate}

@@

%%
CFA2 uses a ``local semantics'' that is similar to our segmenting continuations at function boundaries, but gets stuck when it gets to a point where it would need to ``return.''
%
Returns are un-stuck by an imperative summarization algorithm driving the state exploration.
%
What we show here is almost wholly the same in character, only embodied still in terms of a machine semantics and thus more easily reasoned about, and can be run in the concrete.
%%

@@

\paragraph{Example:} revising the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
%
If the operator were instead $<$ or $+$, we could still constant fold that away.
%%

@@

>>>>

%%
This paper gives a common, simple methodology to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
%
Our presentation is meant to be a tutorial to this related literature, with the additional promise of a novel analysis.
%
We don't presume a language of formal semantics such that our technique applies to all formalisms in such a language.
%
An overly formal approach is unlikely to capture all ``real language's'' formalizations.
%
The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
%
We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
%
The reader unfamiliar with or uninterested in correctness proofs should still be able to grok the technique and apply it with little difficulty.
%
This paper has been partially mechanized in the Coq proof assistant, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.

@@

This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
%
As it is, the CESK machine approximates the tabular machine, but not the other way around;
%
if we use a finite allocation strategy for example, then we don't explore some non-deterministic paths that lead to different results for a function call.
%
We thus explore all paths and share $\mktab$ and $\mmemo$ between all paths.

To turn this semantics into PDCFA's algorithm for constructing a pushdown system with only reachable states, given the appropriate $\alloc$, we apply a widening operator to make $\mktab$, and $\mmemo$ shared amongst all states, in what is then called a $\System$. We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.

\paragraph{Technical aside} There are minor differences with PDCFA's algorithm due to our choice to process one state at a time. If we process the entire frontier in one step, then we need to have additional tables of ``changes to $\mktab$ and $\mmemo$'' that need to be matched up in the following step. Here ``matched up'' means if both tables have the same context in their domain, we must add the state that has the memoized result(s) with the saved continuation(s). If we don't, then some memoized results may not flow to newly reached calling contexts. This corresponds with PDCFA's worksets of edges.
%

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System
\end{align*}
\begin{align*}
  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{align*}
    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
  \end{align*}
}
%%

%%
A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
%
The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
%
We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%
The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
%
Note the ordering on systems is pointwise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
%

\paragraph{Translation from Greek}
 ${\mathcal F}$'s definition may be implemented with destructive updates to the components of $\mathit{sys}$ instead of the big joins and unions.
%
 The definition of $\stepto$ is useful for defining the semantics, but for an implementation we would use a
 function \texttt{step} that takes a state and produces all next states.
%
 We give a completely self-contained implementation in Racket of ${\mathcal F}$ along with \texttt{step} in \autoref{fig:real-impl}.
%
 The only free variable is \texttt{alloc}, which can be instantiated with (for example)
\\[2pt]
 \includegraphics{calloc}
\\
 for concrete allocation, and
\\[2pt]
 \includegraphics{malloc}
\\
 for monovariant allocation.
%
\\[2pt]
\begin{figure}
  \includegraphics[scale=0.8]{impl}
  \caption{Complete Racket implementation}
\label{fig:real-impl}
\end{figure}

\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example.
Note that monovariance means the address for each binding is the variable name itself, so $\menv$ is always the identity function. We therefore omit it.

\begin{center}
  \includegraphics[scale=0.5]{example0}
\end{center}
  % \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))})$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

\begin{figure}
  \centering
      \begin{tabular}{c}
        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
        \\ $\Downarrow$ \\
        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
        \\ $\Downarrow$ \\
        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
      \end{tabular}
      \begin{center}
% KTables and Memo tables
      \begin{tabular}{|l|c|c|c|}
        \hline
        \# & $\mktab$ & $\mmemo$ \\
        \hline
        0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{(0,\mstore_1)}]$ \\
        1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{(0,\mstore_3),(1,\mstore_3)}]$ \\
        \hline
        \end{tabular}
% contexts
      \begin{tabular}{|l|c|}
        \hline
        \# & $\mctx$ \\
        \hline
        0  & $({\tt x}, \mstore_1)$ \\
        1  & $({\tt x}, \mstore_3)$ \\
        \hline
      \end{tabular}
% stores and continuations
      \begin{tabular}{|l|c|c|}
        \hline
        \# & $\mstore$ & $\mkont$ \\
        \hline
        0 & $[{\tt id} \mapsto \set{\slam{\mvar}{\mvar}}]$ & $\mkont_0$ \\
        1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
        2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
        3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
        4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
        \hline
      \end{tabular}
      \end{center}
      Where $\mkont_0 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}}$
      \caption{Example evaluation}
      \label{fig:ex-eval}
\end{figure}
%% FIXME stores in memo-table
\begin{enumerate}
\item{\texttt{(id 0)} steps to \texttt{x} at continuation $\krt{\mctx_0}$ with store $\mstore_1$ and continuation table $\mktab_0$, still with empty memo table.}
\item{\texttt{0} at $\krt{\mctx_0}$ steps to $0$ at $\mkont_0$ and new memo table $\mmemo_0$.}
\item{\texttt{0} at $\mkont_0$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} with store now $\mstore_2$}
\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} with continuation $\mkont_2$.}
\item{\texttt{(id 1)} steps to \texttt{x} with store $\mstore_3$ and and continuation table $\mktab_1$.}
\item{\texttt{0} or \texttt{1} at $\krt{\mctx_1}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ with memo table $\mmemo_1$.}
\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
\end{enumerate}

@@

%%
CFA2 uses a ``local semantics'' that is similar to our segmenting continuations at function boundaries, but gets stuck when it gets to a point where it would need to ``return.''
%
Returns are un-stuck by an external summarization algorithm driving the state exploration.
%
What we show here is almost wholly the same in character, only embodied still in terms of a machine semantics and thus more easily reasoned about, and can be run in the concrete.
%%

@@

\paragraph{Example:} revisiting the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
%
If the operator were instead $<$ or $+$, we could still constant fold that away.
%%

@@

\textbf{TODO:} address FLOPS reviewer's concerns regarding Cook's transformation and Jones' use for analysis.

### 2014-03-25 metatheory.tex
Fixed up (still contrived) example. Added Racket implementation. Small fix to reify and F.

\begin{align*}
  \reify &: \mathit{System} \to \wp(\mathit{State}_\mathit{CESK} \times \mathit{State}_\mathit{CESK}) \\
  \reify(F, R, \mktab, \mmemo) &=
  \setbuild{(\tpl{\mpoint,\mstore,\append{\mkont}{\mkont''}},
             \tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})}
           {\\ &\phantom{= \{}(\tpl{\mpoint,\mstore,\mkont}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
            \\ &\phantom{= \{}\mkont'' \in \tails{\mktab}{\mkont}} \\
\text{where } \tails{\mktab}{\kmt} &= \set{\kmt} \\
              \tails{\mktab}{\krt{\mctx}} &= \setbuild{\mkont' \in \unroll{\mktab}{\mkont}}{\mkont \in \mktab(\mctx)} \\
              \tails{\mktab}{\kcons{\mkframe}{\mkont}} &= \tails{\mktab}{\mkont}
\\[2pt]
              \unroll{\mktab}{\mkont} &= \setbuild{\append{\mkont}{\mkont'}}{\mkont' \in \tails{\mktab}{\mkont}}
\\[2pt]
              \append{\kmt}{\mkont} &= \append{\krt{\mctx}}{\mkont} = \mkont \\
              \append{\kcons{\mkframe}{\mkont}}{\mkont'} &= \kcons{\mkframe}{(\append{\mkont}{\mkont'})}
\end{align*}

The $\mathit{unroll}$ set interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, and $\mathit{tails}$ is defined mutually to produce all continuations that could be unrolled from the tail of a truncated continuation.

@@

>>>>

\begin{align*}
  \reify &: \mathit{System} \to \wp(\mathit{State}_\mathit{CESK} \times \mathit{State}_\mathit{CESK}) \\
  \reify(F, R, \mktab, \mmemo) &=
  \setbuild{(\tpl{\mpoint,\mstore,\append{\mkont}{\mkont''}},
             \tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})}
           {\\ &\phantom{= \{}(\tpl{\mpoint^*,\mstore^*,\mkont^*}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
            \\ &\phantom{= \{}\text{if } \mpoint^*\equiv\mval \text{ and } \mkont^* \equiv \krt{\mctx} \text{ then }
            \\ &\phantom{= \{else }\text{let } (\tpl{\mpoint,\mstore,\mkont},\tpl{\mpoint^*,\mstore^*,\mkont^*}) \in R,
            \\ &\phantom{= \{}\text{else let } \tpl{\mpoint,\mstore,\mkont} = \tpl{\mpoint^*,\mstore^*,\mkont^*}
            \\ &\phantom{= \{}\mkont'' \in \tails{\mktab}{\mkont}} \\
\text{where } \tails{\mktab}{\kmt} &= \set{\kmt} \\
              \tails{\mktab}{\krt{\mctx}} &= \setbuild{\mkont' \in \unroll{\mktab}{\mkont}}{\mkont \in \mktab(\mctx)} \\
              \tails{\mktab}{\kcons{\mkframe}{\mkont}} &= \tails{\mktab}{\mkont}
\\[2pt]
              \unroll{\mktab}{\mkont} &= \setbuild{\append{\mkont}{\mkont'}}{\mkont' \in \tails{\mktab}{\mkont}}
\\[2pt]
              \append{\kmt}{\mkont} &= \append{\krt{\mctx}}{\mkont} = \mkont \\
              \append{\kcons{\mkframe}{\mkont}}{\mkont'} &= \kcons{\mkframe}{(\append{\mkont}{\mkont'})}
\end{align*}

Any return edge (\ie, the left state has a value and a $\krt{\mctx}$ continuation) is administrative, so the reification stutters those out by linking the predecessors of return states with their successors.
%
The $\mathit{unroll}$ set interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, and $\mathit{tails}$ is defined mutually to produce all continuations that could be unrolled from the tail of a truncated continuation.

@@

To reiterate, this correctness proof is strong enough for any allocation strategy, including fresh (for concrete execution) and finite (for ``pushdown analysis'').
%
The finite case is special since the $\System$ space becomes finite, which means there is some $m$ such that ${\mathcal F}^m(\mastate_0)$ is a fixed point, by Kleene's fixed point theorem.
%
In other words, the fixed point soundly and completely represents all paths reachable from the initial state in the CESK machine with an unbounded stack but finite address space.
%
This form of the CESK machine has a shallow translation into a pushdown system, but we needn't go there.
